### 2018-09-23 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\cite{Yang2017} MidiNet
Following this light, we investigate in this paper a novel CNN-based model for symbolic-domain generation, focus- ing on melody generation. 1 Instead of creating a melody sequence continuously, we propose to generate melodies one bar (measure) after another, in a successive manner. This allows us to employ convolutions on a 2-D matrix representing the presence of notes over different time steps in a bar. We can have such a score-like representation for each bar for either a real or a generated MIDI.Moreover, to emulate creativity [23] and encourage diverse generation result, we use random noises as input to our generator CNN. The goal of the generator is to trans- form random noises into the aforementioned 2-D score- like representation, that “appears” to be from real MIDI. This transformation is achieved by a special convolution operator called transposed convolution [8]. Meanwhile, we learn a discriminator CNN that takes as input a 2-D score- like representation and predicts whether it is from a real or a generated MIDI, thereby informing the generator how to appear to be real. This amounts to a generative adversarial network (GAN) [11–13,24,27], which learns the generator and discriminator iteratively under the concept of minimax two-player game theory. This GAN alone does not take into account the temporal dependencies across different bars. To address this issue, we propose a novel conditional mechanism to use music from the previous bars to condition the generation of the present bar. This is achieved by learning another CNN model, which we call the conditioner CNN, to incorporate information from previous bars to intermediate layers of the generator CNN. This way, our model can “look back” without a recurrent unit as used in RNNs. Like RNNs, our model can generate music of arbitrary number of bars.
For simplicity, we filtered out MIDI tabs that contain chords other than the 24 basic chord triads (12 major and 12 minor chords). Next, we segmented the remaining tabs every 8 bars, and then pre-processed the melody channel and the chord channel separately, as described below. For melodies, we fixed the smallest note unit to be the sixteenth note, makingw = 16. Specifically, we prolonged notes which have a pause note after them. If the first note of a bar is a pause, we extended the second note to have it played while the bar begins. There are other exceptions such as triplets and shorter notes (e.g. 32nd notes), but we chose to exclude them in this implementation. More- over, for simplicity, we shifted all the melodies into two oc- taves, from C4 to B5, and neglected the velocity of the note events. Although our melodies would use only 24 possible notes after these preprocessing steps, we considered all the 128 MIDI notes (i.e. from C0 to G10) in our symbolic representation. In doing so, we can detect model collaps- ing [12] more easily, by checking whether the model gen- erates notes outside these octaves. As there are no pauses in our data after preprocessing, we do not need a dimension for silence. Therefore, h = 128. For chords, instead of using a 24-dimensional one-hot vector, we found it more efficient to use a chord representa- tion that has only 13 dimensions— the first 12 dimensions for marking the key, and the last for the chord type (i.e. major or minor), as illustrated in Table 2. We pruned the chords such that there is only one chord per bar. After these preprocessing steps, we were left with 526 MIDI tabs (i.e. 4,208 bars). 5 For data augmentation, we circularly shifted the melodies and chords to any of the 12 keys in equal temperament, leading to a final dataset of 50,496 bars of melody and chord pairs for training.
Because we use random noises as inputs to our generator, our model can generate melodies from scratch, i.e. without any other prior information. However, due to the conditioner CNN, our model has the capacity to exploit whatever prior knowledge that is available and can be rep- resented as a matrix. For example, our model can generate music by following a chord progression, or by following a few starting notes (i.e. a priming melody). Given the same priming melody, our model can generate different results each time, again due to the random input.
Symbolic Representation for Convolution
Our model uses a symbolic representation of music in fixed time length, by dividing a MIDI file into bars. The note events of a MIDI channel can be represented by an h- by-w real-valued matrix X, where h denotes the number of MIDI notes we consider, possibly including one more dimension for representing silence, and w represents the number of time steps we use in a bar. For melody gener- ation, there is at most one active note per time step. We use a binary matrixX ∈ {0, 1}h×w if we omit the velocity (volume) of the note events. We use multiple matrices per bar if we want to generate multi-track music. In this representation, we may not be able to easily distinguish between a long note and two short repeating notes (i.e. consecutive notes with the same pitch). Future exten- sions can be done to emphasize the note onsets.
Generator CNN and Discriminator CNN
The core of MidiNet is a modified deep convolutional gen- erative adversarial network (DCGAN) [24], which aims at learning a discriminatorD to distinguish between real (au- thentic) and generated (artificial) data, and a generator G that “fools” the discriminator. Our discriminator is a typical CNN with a few convolution layers, followed by fully-connected layers. These lay- ers are optimized with a cross-entropy loss function, such that the output of D is close to 1 for real data (i.e. X) and 0 for those generated (i.e. G(z)). We use a sigmoid neuron at the output layer of D so its output is in [0,1].
Conditioner CNN
We propose to achieve this by using a conditioner CNN
that can be viewed as a reverse of the generator CNN. As the blue blocks in Figure 1 illustrates, the conditionerCNN uses a few convolution layers to process the input h-by-w conditional matrix. The conditioner and generator CNNs use exactly the same filter shapes in their convolution lay- ers, so that the outputs of their convolution layers have “compatible” shapes. In this way, we can concatenate the output of a convolution layer of the conditioner CNN to the input of a corresponding transposed convolution layer of the generator CNN, to influence the generation process. In the training stage, the conditioner and generator CNNs are trained simultaneously, by sharing the same gradients.
Tunning for Creativity
We propose two methods to control the trade-off between creativity and discipline of MidiNet. The first method is to manipulate the effect of the conditions by using them only in part of the intermediate transposed convolution layers of G, to give G more freedom from the imposed conditions. The second method capitalizes the effect of the fea- ture matching technique [27]: we can increase the values of λ1 and 2 to make the generated music sounds closer to existing music (i.e. those observed in the training set).
As the major task considered in this paper is melody generation, for training MidiNet we need a MIDI dataset that clearly specifies per file which channel corresponds to the melody. To this end, we crawled a collection of 1,022 MIDI tabs of pop music from TheoryTab, 4 which provides exactly two channels per tab, one for melody and the other for the underlying chord progression. With this dataset, we can implement at least two versions of MidiNets: one that learns from only the melody channel for fair comparison with MelodyRNN [33], which does not use chords, and the other that additionally uses chords to condition melody generation, to test the capacity of MidiNet.
Network Specification
Our model was implemented in TensorFlow. For the gen- erator, we used as input random vectors of white Gaussian noise of length l = 100. Each random vector go through two fully-connected layers, with 1024 and 512 neurons re- spectively, before being reshaped into a 1-by-2 matrix. We then used four transposed convolution layers: the first three use filters of shape 1-by-2 and two strides [8], and the last layer uses filters of shape 128-by-1 and one stride. Accord- ingly, our conditioner has four convolution layers, which use 128-by-1 filters for the first layer, and 1-by-2 filters for the other three. For creating a monophonic note sequence, we added a layer to the end of G to turn off per time step all but the note with the highest activation.

To evaluate the aesthetic quality of the generation result, a user study that involves human listeners is needed. We conducted a study with 21 participants. Ten of them un- derstand basic music theory and have the experience of be- ing an amateur musician, so we considered them as people with musical backgrounds, or professionals for short. We compared MidiNet with three MelodyRNN mod-
els pre-trained and released by Google Magenta: the basic RNN, the lookback RNN, and the attention RNN [33]. We randomly picked 100 priming melodies from the training data 7 and asked the models create melodies of eight bars by following these primers. We considered two variants of MidiNet in the user study: model 1 (Section 4.2.1) for fair comparison with MelodyRNN, and model 2 (Section 4.2.2) for probing the effects of using chords. Although the result of model 2 was generated by additionally following the chords, we did not playback the chord channel in the user study. We randomly selected the generation result of three out
of the 100 priming melodies for each participant to listen to, leading to three sets of music. To avoid bias, we ran- domly shuffled the generation result by the five considered models, such that in each set the ordering of the five mod- els is different. The participants were asked to stay in a quiet room separately and used a headphone for music lis- tening through the Internet, one set at a time. We told them that some of the music “might be” real, and some might be generated by machine, although all of them were actu- ally automatically generated. They were asked to rate the generated melodies in terms of the following three metrics: how pleasing, how real, and how interesting, from 1 (low) to 5 (high) in a five-point Likert scale.

@@

\section{Податочна репрезентација }

@@

\chapter{Архитектури}

\section{Едноставен ЛСТМ модел}

\section{Проширен ЛСТМ модел}

\section{Енкодер-Декодер модел}

@@

Моделот за учење се состои 3 подмодели, по еден за секој од инструментите. Моделите меѓусебе се идентични, претставени со (ЛИНК ДО ДИЈАГРАМ НА МОДЕЛ). Една влезна рамка за секој од моделите се сотоид од 3 вектори кои ја претставуваат тековно исвирената комбинација на ноти за конкретниот инструмент во временскиот момент кој го претставува рамката. Моделите започнуваат со 3 блока, еден блок кој на влез прима N временски рамки од тековниот временски момент наназад, еден блок на влез ја примат тековната временска рамка, а последниот модел прима N временски рамки нанапред.

@@

Двете LSTM мрежи се составени од два слоја по 256 LSTM јазли. Влезната димензија е еднаква на излезот од ембединг слојот помножен со бројот на временски чекори кој се гледа наназад, или нанапред (32), познат како lookback, а излезот е реалновредностен вектор со 256 елементи, со активациска функција тангенс хуперболикус. Излезите ги претставуваат предвидувањата за тековниот чекор врз основа на временскиот сегмент што го надгледува конкретниот блок.

@@

@@

\subsection{Паралелно Гибсово семплирање}

\subsection{Температура}

>>>>

Многу интересен концепт во полето на машинско учење претставуваат Генеративните Непријателски Мрежи (анг. Generative Adverserial Networks - GAN) каде што две подмрежи работат енда против друга. Првата мрежа, наречена "генератор", учи како да генерира лажни примероци врз основа на оригиналното множество со цел да ја излаже втората мрежа, наречена "дискриминатор", чија задача е да одреди дали примерокот кој и е прикажан е реален припаник на оригиналното податочно множество или не. Добро истрениран генератор може да се искористи за генеирање на музички секвенци. Во \cite{Yang2017,Dong2017,Dong2018} се пиркажани имплементации на GAN системи со користење на конволуциски мрежи во генераторот и дискриминаторот. Конволуциските мрежи третираат рамки од еден или повеќе такта како поединечни влезно/излезни елементи. Во вакв форма мрежите може да научат зависности само помеѓу нотите во рамки на такт, а не на подолг период, или пак помеѓу тактови. За да го решат овој проблем аврорите во Во \cite{Yang2017} користат паралелна конволуциска мрежа за условување на генераторот, која му дава на влез условена верзија од претходниот временски чекор на генераторот, со цел учење на временски зависности помеѓи два такта. Во трудовите \cite{Dong2017,Dong2018} се прикажани поамбициозни модели за генерирање на повеќе музички траки паралелно со користење на GAN систем со конволуциски генератор и дискриминатор. Секоја музичка трака има свој подмодел, кој почнува со дел за моделирање/генерирање на компонента која ја опишува песната на повисоко ниво (сублимира секвенци од тактови), која потоа служи како еден од влезовите во остатокот од системот кој што учи/генерира цели тактови. Во \cite{Dong2018} авторите додават на излез од системо додатна мрежа наменета за подобрување на реконструкцијата на тактови од реално вредносните вектори кои се излез од конволуциските мрежи, и тие се структурирани како DBN.

@@

@@

@@

Архитектурата е имплементирана со користење на Keras библиотеката за длабоко учење, па така некои термини и имиња доаѓаат од терминологијата на библиотеката.
Моделот за учење се состои 3 подмодели, по еден за секој од инструментите. Моделите меѓусебе се идентични, претставени со (ЛИНК ДО ДИЈАГРАМ НА МОДЕЛ). Една влезна рамка за секој од моделите се сотоид од 3 вектори кои ја претставуваат тековно исвирената комбинација на ноти за конкретниот инструмент во временскиот момент кој го претставува рамката. Моделите започнуваат со 3 блока, еден блок кој на влез прима N временски рамки од тековниот временски момент наназад, еден блок на влез ја примат тековната временска рамка, а последниот модел прима N временски рамки нанапред.

@@

Двете LSTM мрежи се составени од два слоја по 256 LSTM јазли. Влезната димензија е еднаква на излезот од ембединг слојот помножен со бројот на временски чекори кој се гледа наназад, или нанапред (32), познат како lookback, а излезот е реалновредностен вектор со 256 елементи, со активациска функција тангенс хуперболикус. Излезите ги претставуваат предвидувањата за тековниот чекор врз основа на временскиот сегмент што го надгледува конкретниот блок. Бројот на временски чекори кој се гледа наназад и или нанапред е всушност Back Propagation Through Time, или поточно неговата скратена варијанта Truncated Back Propagation, во Keras одредена со големината на субсеквенца од временски чекори коишто и се даваат на LSTM подмрежата во одреден момент. 
(ДИЈАГРАМ ОД СЛОЈ)

@@

\section{Тренрање на моделот}

@@

Генерирање на песните го вршам со Псеудо-Гибсова техника за вадење на примероци од веројатносна дистрибуција, заедно со ситем за паралелно ажурирање на вредности (ЦИТИРАЈ PARALLEL GIBBS). Генерирањето започнува со симболот за почеток и трае до предвидување на симбол за крај или до максимална зададена должина на секвенца. Процесот на генерирање е итеративен, т.е. секој претходно предвиден временски чекор влијае на следното предвидување. Техниката предвидува употреба на параметар температура, кој што воведува стохастичност во процесот. Искористив почетна вредност од 1.5 и минимална вредност од 1.0. Колку поголема вредноста има температурата толку е понепредвидлив изборот на следниот елемент во процесот на предвидување, а колку е понизок процесот станува се по-детерминистички.
